---
title: "a3_task4_mikel_irigoyen.Rmd"
author: "Mikel irigoyen"
date: "11/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = FALSE)
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
```
## In this section wecompare the mean size of purple urchins in CONTROL (no giant kelp removal) or ANNUAL (annual kelp removal) sections of Naples Reef in January 2010. 

- here i read in the data and clean_names
```{r}
urchins <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names()
```
###Wrangling data to only include Naples Reef observations for purple urchins observed in January 2010

- Included some previously used code from task 3 

- Began by using lubridate to format dates 

- Then used case_when() to refer to NAPL by its full name in new column

- After that used case_when() to create annual and control in its own column

- Then used filter() to only see Naples observations for the month/year that we are looking for.
```{r}
naples_obs <- urchins %>% 
  
 mutate(date = mdy(date), 
        month = month(date),  
        year = year(date),
        site_full = case_when(site == "NAPL" ~ "Naples Reef")) %>%
  mutate(treatment = case_when(treatment == "ANNUAL" ~ "annual", treatment == "CONTROL" ~ "control")) %>% 
  filter(common_name == "Purple Urchin", site == "NAPL",
         month == 1, year == 2010)

  
```
## Histograms and QQ-plots

- Using these to look at the size distribution for each of the treatments (Control and Annual) in order to assess normality

- First up is the histogram, used facet_wrap() to get separate plots for both treatment types
```{r}
ggplot(naples_obs, aes(size)) +
  geom_histogram()+
  facet_wrap(~treatment)
```

- Next we do the qq-plot

- Same procedure that was used on the histogram to get results

- One difference being, qq-plots require knowing a sample, so must include "sample = size" 

```{r}
ggplot(naples_obs, aes(sample = size))+
  geom_qq()+
  facet_wrap(~treatment)
```

- The histograms show a relatively normal distribution and the linear trend in the qq-plots do as well. This shows normality between the two types of urchins.

## Summary Statistics Time!!!
- Here we create a basic table of descriptive statistics for both groups that will include the mean, standard deviation, and sample size

```{r}
purp_urch_stats <- naples_obs %>% 
  group_by(treatment) %>% 
  summarize(naples_mean = mean(size),
            naples_sd = sd(size),
            count = n())
print(purp_urch_stats)
```

## T-test time
- Here we perform a two-sample t-test to compare means between the two groups

- Created separate vectors differentiated by "treatment"

- Pull by "size"

- Then do a t-test with annual vs. control

```{r}
ann_urchins <- naples_obs %>% 
  filter(treatment == "annual") %>% 
  pull(size)
```

```{r}
con_urchins <- naples_obs %>% 
  filter(treatment == "control") %>% 
  pull(size)
```
```{r}
t.test(x = ann_urchins, y = con_urchins)
```
Between annual urchins(3.91667) and control urchins(3.725806) we see a mean % difference of 4.99482%. This low percentage should suggest they come from the same population. With such a large p-value(0.2536) we are to reject the null hypothesis and this means there is a 25% chance of getting means that are at least this different, if they are drawn from urchin populations with the same mean.
